### 2025-08-18 chapters/background.tex
5.3 complete

@@

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.9\linewidth]{figures/keystone_overview.png}
\caption{Overview of the Keystone architecture illustrating components such as the Security Monitor, enclave runtime, and the privilege hierarchy.}
\label{fig:keystone_overview}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.9\linewidth]{figures/enclave_lifecycle.png}
\caption{Stages of a Keystone enclave lifecycle: creation, execution, and destruction.}
\label{fig:enclave_lifecycle}
\end{figure}

@@

@@

@@

@@

@@

@@

Keystone effectively synthesizes the strengths of these platforms while avoiding their limitations. It supports dynamic enclave creation spanning both user and supervisor modes, with security ensured via low-overhead Physical Memory Protection (PMP) and a lightweight, verifiable Security Monitor operating in M-mode \cite{dayeol2019keystone}. This design offers richer enclave functionality with reduced complexity.

@@

@@

@@

@@

>>>>

\begin{figure}[htbp]
\centering
\begin{subfigure}[b]{0.48\linewidth}
  \centering
  \begin{tikzpicture}[font=\small]
    \draw[fill=green!10] (0,2) rectangle (3,5) node[midway]{\textbf{Secure World}};
    \draw[fill=red!10] (5,2) rectangle (8,5) node[midway]{\textbf{Normal World}};
    \draw[fill=gray!20] (2,0) rectangle (6,1.5) node[midway]{\textbf{Secure Monitor}};
    \draw[->,thick] (3,3.5) -- (5,3.5) node[midway,above]{\textbf{SMC}};
  \end{tikzpicture}
  \caption{TrustZone's dual-world architecture with SMC-managed transitions.}
  \label{fig:trustzone}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.48\linewidth}
  \centering
  \begin{tikzpicture}[font=\small]
    \node[above] at (3,4) {\textbf{Enclave Page Cache (EPC)}};
    \draw[thick] (0,0) rectangle (6,4);

    \foreach \y/\label in {3.0/{Page 1 (Enc)}, 2.0/{Page 2 (Enc)}, 1.0/{Page 3 (Enc)}} {
      \fill[blue!10] (0.3,\y) rectangle (5.7,\y+0.8);
      \draw (0.3,\y) rectangle (5.7,\y+0.8) node[midway]{\label};
    }
  \end{tikzpicture}
  \caption{Intel SGX enclaves mapped to encrypted pages inside the EPC.}
  \label{fig:sgx-epc}
\end{subfigure}
\caption{Comparison of two TEE memory architectures: (a) ARM TrustZone split-world and (b) Intel SGX EPC memory model.}
\label{fig:tee-architectures}
\end{figure}

@@

\begin{table}[htbp]
\centering
\small
%\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}p{3cm}p{2.5cm}p{2.5cm}p{2.5cm}@{}}
\toprule
\textbf{Feature} & \textbf{Intel SGX} & \textbf{ARM TrustZone} & \textbf{Keystone} \\
\midrule
Privilege Model & Ring 3 enclaves & Secure Monitor & M-mode TCB \\
Openness & Closed & Closed & Open \\
Use Cases & Cloud security & Mobile security & Research, IoT \\
Hardware Dependency & Intel CPUs & ARM SoCs & RISC-V \\
\bottomrule
\end{tabular}
\caption{Comparison of three TEE implementations.}
\end{table}

@@

\begin{figure}[htbp]
\centering
\begin{subfigure}[b]{0.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/keystone_overview.png}
    \caption{Keystone architecture showing components such as the Security Monitor, enclave runtime, and the privilege hierarchy.}
    \label{fig:keystone_overview}
\end{subfigure}

\vspace{0.5cm}

\begin{subfigure}[b]{0.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/enclave_lifecycle.png}
    \caption{Lifecycle of a Keystone enclave: creation, execution, and destruction phases.}
    \label{fig:enclave_lifecycle}
\end{subfigure}

\caption{Overview of the Keystone architecture and its enclave lifecycle.}
\label{fig:keystone_grouped}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[node distance=2.5cm,>=Stealth]
\tikzstyle{block} = [draw, minimum width=3cm, minimum height=2cm, align=center, font=\small]

\node[block, fill=gray!10] (cpu) {CPU Core\\\footnotesize Modes: M / S / U};
\node[block, fill=gray!20, right=of cpu] (pmp) {PMP Unit};
\node[block, fill=gray!30, right=of pmp] (mem) {Physical Memory};

\draw[->, thick] (cpu.east) -- node[above]{\footnotesize Access Request} (pmp.west);
\draw[->, thick] (pmp.east) -- node[above]{\footnotesize Permission Check} (mem.west);
\end{tikzpicture}
\caption{PMP hardware architecture in the RISC‑V privilege hierarchy. Memory accesses from the CPU are filtered through PMP entries to enforce physical memory protection.}
\label{fig:pmp-arch}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[node distance=2.8cm,>=Stealth]
\usetikzlibrary{positioning, backgrounds}
\tikzstyle{block} = [draw, minimum width=3cm, minimum height=1.5cm, font=\small]

% Main nodes
\node[block, fill=gray!10] (os) {OS (S-mode)};
\node[block, fill=gray!20, right=of os] (pmpunit) {PMP Unit};
\node[block, fill=gray!30, right=of pmpunit] (enclave) {Enclave (EPM)};

% Arrows
\draw[->, thick] (os.east) -- node[above]{\footnotesize Memory Access} (pmpunit.west);
\draw[->, thick] (pmpunit.east) -- node[above]{\footnotesize Filtered Access} (enclave.west);

% Context switch label (moved lower)
\node[below=1.8cm of pmpunit, font=\footnotesize, fill=white, inner sep=2pt, rounded corners=1pt, draw=gray!50] 
    {Context switch $\Rightarrow$ PMP reconfig};

% Background shading (after everything else)
\begin{scope}[on background layer]
  % Only OS is untrusted
  \fill[red!10, opacity=0.3] 
    ($(os.north west)+(-0.4,0.4)$) rectangle 
    ($(os.south east)+(0.4,-0.4)$);

  % PMP + Enclave are trusted
  \fill[green!10, opacity=0.3] 
    ($(pmpunit.north west)+(-0.4,0.4)$) rectangle 
    ($(enclave.south east)+(0.4,-0.4)$);
\end{scope}

% Trust domain labels (moved closer to diagram)
\node[below=0.8cm of os, font=\scriptsize, align=center, fill=white, inner sep=2pt, draw=gray!40, rounded corners=1pt] 
    {Untrusted domain};

\node[below=0.8cm of enclave, font=\scriptsize, align=center, fill=white, inner sep=2pt, draw=gray!40, rounded corners=1pt] 
    {Trusted enclave memory};

\end{tikzpicture}
\caption{PMP-enforced isolation of Keystone enclaves. PMP is reconfigured on each context switch to protect enclave memory from the untrusted OS.}
\label{fig:pmp-keystone}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
% Memory line
\draw[thick] (0,0) -- (12,0);

% Address labels
\foreach \x/\lbl in {0/0x0000, 4/0x4000, 8/0x8000, 12/0xC000} {
    \draw (\x,0.1) -- (\x,-0.1) node[below, font=\scriptsize] {\lbl};
}

% Entry regions
\draw[fill=blue!30, opacity=0.8] (2,0.3) rectangle (8,0.8)
    node[pos=.5, font=\footnotesize] {Entry 1};
\draw[fill=red!40, opacity=0.6] (4,0.3) rectangle (10,0.8)
    node[pos=.5, font=\footnotesize, text=black] {Entry 0};

% Priority indication
\draw[->, thick] (10.5,0.55) -- (11.5,0.55) node[right, font=\scriptsize] {Higher Address};
\node[below, font=\footnotesize] at (6,-0.3) {Lower-numbered entry (Entry 0) takes precedence over Entry 1};
\end{tikzpicture}
\caption{Overlapping PMP regions. When regions overlap, the lower-numbered PMP entry has priority.}
\label{fig:pmp-priority}
\end{figure}

@@

\begin{table}[htbp]
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{>{\raggedright}p{3cm} >{\raggedright}p{4cm} >{\raggedright\arraybackslash}p{5cm}}
\hline
\textbf{Limitation} & \textbf{Impact} & \textbf{Enhancement / Research} \\
\hline
Fixed \# entries & Limits scalability & Cerberus verified sharing \\
No dynamic resizing & Hard to adapt to workloads & Dynamic PMP proposals \\
Limited granularity & Coarse protection & Fine‑grained tagging \\
\hline
\end{tabular}
\caption{Summary of PMP limitations and research‑driven enhancements.}
\label{tab:pmp-limitations}
\end{table}

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
    xbar,
    width=10cm, height=5cm,
    xmin=0, xmax=100,
    xlabel={Relative TCB size (\%)},
    symbolic y coords={SEV,SGX,TrustZone,Keystone},
    ytick=data,
    nodes near coords,
    every node near coord/.append style={font=\scriptsize},
    bar width=10pt
]
\addplot[pattern=north east lines] coordinates {(60,SEV) (50,SGX) (80,TrustZone) (20,Keystone)};
\end{axis}
\end{tikzpicture}
\caption{Estimated relative TCB sizes (illustrative).}
\label{fig:tcb-size}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={TrustZone SMC,SGX ECALL,Keystone call},
    xtick=data,
    ylabel={Latency ($\mu$s)},
    ymin=0, ymax=12,
    nodes near coords,
    bar width=14pt
]
\addplot[pattern=dots] coordinates {(TrustZone SMC,10) (SGX ECALL,8) (Keystone call,3)};
\end{axis}
\end{tikzpicture}
\caption{Illustrative context switch latency across TEEs.}
\label{fig:ctx-overhead}
\end{figure}
Keystone effectively synthesizes the strengths of these platforms while avoiding their limitations. It supports dynamic enclave creation spanning both user and supervisor modes, with security ensured via low-overhead Physical Memory Protection (PMP) and a lightweight, verifiable Security Monitor operating in M-mode \cite{dayeol2019keystone}. This design offers richer enclave functionality with reduced complexity.

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
    ybar,
    bar width=8pt,
    width=10cm, height=6cm,
    ymin=0, ymax=110,
    ylabel={Relative performance (\%)},
    symbolic x coords={CPU-bound,Memory-bound,I/O-bound},
    xtick=data,
    ymajorgrids,
    legend style={at={(0.5,-0.2)},anchor=north,legend columns=3}
]
\addplot[pattern=north east lines] coordinates {(CPU-bound,98) (Memory-bound,70) (I/O-bound,65)};
\addplot[pattern=dots] coordinates {(CPU-bound,100) (Memory-bound,50) (I/O-bound,55)};
\addplot[pattern=crosshatch] coordinates {(CPU-bound,95) (Memory-bound,60) (I/O-bound,50)};
\legend{Keystone,SGX,TrustZone}
\end{axis}
\end{tikzpicture}
\caption{Illustrative TS‑Perf‑style benchmark comparison (REE=100\%).}
\label{fig:bench-compare}
\end{figure}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[>=stealth, font=\small]
\draw[->, thick] (0,0) -- (12,0);

\foreach \x/\year/\name in {0/2004/TrustZone,3/2015/SGX,5/2016/SEV,8/2019/Keystone,10/2020/TDX} {
    \draw (\x,0) -- (\x,-0.3) node[below] {\year};
    \node[above=0.3cm] at (\x,0) {\name};
}
\end{tikzpicture}
\caption{Timeline of selected TEE introductions.}
\label{fig:tee-timeline}
\end{figure}

@@

\begin{table}[htbp]
\centering
\caption{Feature comparison of major TEEs.}
\label{tab:tee-features}
\begin{tabular}{@{}lcccc@{}}
\toprule
Feature & Keystone & SGX & TrustZone & SEV \\
\midrule
Open Source & Yes & No & No & No \\
Dynamic Enclaves & Yes & No & Limited & No \\
Hardware Dependency & RISC-V & Intel & ARM & AMD \\
TCB Size & Low & Medium & High & Medium \\
Flexible Configurations & High & Low & Low & Medium \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
    width=10cm, height=6cm,
    xlabel={Working set (MB)},
    ylabel={Relative performance (\%)},
    xmin=64, xmax=1024,
    ymin=0, ymax=110,
    xtick={64,128,256,512,768,1024},
    ytick={0,20,40,60,80,100},
    xmajorgrids, ymajorgrids, grid style=dashed,
    ticklabel style={font=\small}
]
% Shaded region beyond EPC size (example: 256 MB)
\path[fill=red!12, draw=none] (axis cs:256,0) rectangle (axis cs:1024,110);

% Performance curve (illustrative)
\addplot[black, very thick, mark=*, mark options={fill=black}] coordinates {
(64,100) (128,98) (256,95) (512,70) (768,40) (1024,20)
};

% EPC boundary
\addplot[dashed, black, thick] coordinates {(256,0) (256,110)};
\node[font=\footnotesize, fill=white, inner sep=1pt, anchor=south east]
    at (axis cs:256,110) {EPC boundary};

\end{axis}
\end{tikzpicture}
\caption{Performance drops sharply once the working set exceeds the protected memory size (shaded)}
\label{fig:epc-cliff}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
    ybar,
    bar width=12pt,
    width=10cm, height=6cm,
    ymin=0, ymax=110,
    ylabel={Relative throughput (\%)},
    symbolic x coords={Write,Read,Rewrite},
    xtick=data,
    ymajorgrids, grid style=dashed,
    legend style={at={(0.5,-0.25)},anchor=north,legend columns=2, font=\small},
    pattern color=black,
    nodes near coords,
    every node near coord/.append style={font=\scriptsize}
]
% Enclave (throughput = 100 - slowdown)
\addplot[pattern=north east lines] coordinates {(Write,63.8) (Read,59.1) (Rewrite,44.9)};
% REE baseline
\addplot[pattern=vertical lines] coordinates {(Write,100) (Read,100) (Rewrite,100)};
\legend{Enclave,REE}
\end{axis}
\end{tikzpicture}
\caption{I/O throughput degradation in enclaves relative to REE: write (−36.2\%), read (−40.9\%), and record rewrite (−55.1\%).}
\label{fig:io-throughput}
\end{figure}

### 2025-08-18 chapters/benchmark.tex
5.3 complete

rkloads, \texttt{coremark} stood out as the only one to exhibit consistent, positive scaling behavior. Performance improved by approximately 40.8\% with each additional core, and a similarly proportional uplift was observed with increased memory. This result is aligned with earlier observations: CoreMark’s mixed-instruction profile and parallelism-friendly design allow it to exploit both increased computational throughput and expanded memory capacity. As a compute-bound benchmark with moderate data locality, CoreMark scales predictably and efficiently with additional hardware provisioning.

@@

\section{Impact of Hardware Parameters (Cores, Memory, Cache)}
\label{sec:hardware-impact}

Following the detailed analysis of enclave overheads and workload-specific sensitivities, this section investigates how changes in underlying hardware parameters—specifically CPU core count and available memory—affect performance across a representative suite of benchmarks. Using the same low-end RISC-V platform and execution model, we systematically varied hardware configurations by incrementally increasing CPU cores and memory capacity. This approach allows us to empirically evaluate whether augmenting computational or memory resources consistently improves performance or if it leads to inefficiencies or stagnation.

We evaluated six workloads, including both native and enclave variants: \texttt{coremark}, \texttt{coremark.ke}, \texttt{dhrystone}, \texttt{enclave.ke}, \texttt{kyber}, and \texttt{kyber.ke}. Each workload was tested across several hardware configurations to characterize scaling behavior. Table~\ref{tab:hardware-impact} summarizes the number of workloads that showed improvement, slowdown, or no change when hardware resources were increased. The table also reports average relative performance changes per incremental scaling step.

@@

Despite the intuitive expectation that adding more cores or memory would improve performance, the results reveal a nuanced and heterogeneous landscape. Only one-third of workloads exhibit performance gains, while half remain largely unaffected. Notably, \texttt{coremark} and its enclave variant are the primary beneficiaries of increased resources. Conversely, \texttt{dhrystone} and \texttt{kyber} workloads show degradation or no meaningful scaling, potentially due to workload characteristics or overheads associated with enclave execution and system-level bottlenecks.

To rigorously assess statistical significance, we conducted both parametric (ANOVA) and non-parametric (Kruskal-Wallis) tests on the performance data (Figure~\ref{fig:anova-kruskal-5pattern-wide-labels}). Across all workloads and hardware parameters, none of the observed performance changes reached significance at the $\alpha = 0.05$ level. \texttt{coremark} and \texttt{kyber} showed marginal p-values close to the threshold in both tests, but their effect sizes remained weak. This outcome was consistent across both test families, affirming that the observed variability in performance is unlikely to result from the number of cores or memory alone.\footnote{Kruskal-Wallis tests are used here as a non-parametric alternative to ANOVA to account for violations of normality and homoscedasticity assumptions.}

% \begin{table}[htbp]
% \centering
% \begin{threeparttable}
% \caption{ANOVA and Kruskal-Wallis results for hardware scaling effects. No statistically significant relationships found ($\alpha=0.05$).}
% \label{tab:anova-corr}
% \begin{tabular}{@{}l
%                 S[table-format=1.2] S[table-format=1.3]
%                 S[table-format=1.2] S[table-format=1.3]
%                 S[table-format=1.2] S[table-format=1.3]
%                 S[table-format=1.2] S[table-format=1.3]
%                 S[table-format=1.3]@{}}
% \toprule
% \textbf{Workload} 
% & \textbf{F (C)} & \textbf{p (C)} 
% & \textbf{F (M)} & \textbf{p (M)} 
% & \textbf{H (C)} & \textbf{p (C)} 
% & \textbf{H (M)} & \textbf{p (M)} 
% & \textbf{$R^2$} \\
% \midrule
% \texttt{coremark}      & 1.50 & 0.218 & 1.50 & 0.218 & 4.60 & 0.100 & 4.74 & 0.094 & 0.030 \\
% \texttt{coremark.ke}   & 0.01 & 0.975 & 0.01 & 0.975 & 0.02 & 0.989 & 0.00 & 1.000 & 0.000 \\
% \texttt{dhrystone}     & 0.28 & 0.837 & 0.28 & 0.837 & 0.71 & 0.870 & 0.65 & 0.883 & 0.005 \\
% \texttt{enclave.ke}    & 0.01 & 0.998 & 0.01 & 0.998 & 0.00 & 1.000 & 0.00 & 1.000 & 0.000 \\
% \texttt{kyber}         & 1.56 & 0.202 & 1.56 & 0.202 & 4.71 & 0.094 & 4.52 & 0.104 & 0.036 \\
% \texttt{kyber.ke}      & 0.00 & 1.000 & 0.00 & 1.000 & 0.00 & 1.000 & 0.00 & 1.000 & 0.000 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes}
% \footnotesize
% \item F: ANOVA F-statistic; H: Kruskal-Wallis H-statistic; p: p-value; $R^2$: coefficient of determination.
% \item (C): CPU core scaling; (M): memory scaling.
% \item Kruskal-Wallis used due to potential non-normality in small-sample performance data\footnote{Kruskal-Wallis is robust to non-normal distributions and unequal variances.}.
% \end{tablenotes}
% \end{threeparttable}
% \end{table}

\begin{figure}[htbp]
\centering
\begin{threeparttable}
    \begin{tikzpicture}

    \begin{axis}[
        ybar,
        bar width=8pt,
        width=14cm,
        height=8cm,
        ymin=0,
        ymax=5.0,
        ylabel={F / H statistic value},
        symbolic x coords={coremark,coremark.ke,dhrystone,enclave.ke,kyber,kyber.ke},
        xtick=data,
        enlarge x limits=0.15,
        ymajorgrids,
        grid style=dashed,
        legend style={at={(0.5,-0.28)},anchor=north,legend columns=3},
        axis y line*=left,
        axis x line*=bottom,
        x tick label style={rotate=45,anchor=east},
        nodes near coords,
        point meta=explicit symbolic,
        every node near coord/.append style={rotate=90, anchor=west, font=\footnotesize}
    ]

    % F (C) - dots
    \addplot[pattern=dots,pattern color=black] 
    coordinates {
    (coremark,1.50)   [1.50]
    (coremark.ke,0.01) []
    (dhrystone,0.28)  [0.28]
    (enclave.ke,0.01) []
    (kyber,1.56)      [1.56]
    (kyber.ke,0.00)   []
    };

    % F (M) - horizontal lines
    \addplot[pattern=horizontal lines,pattern color=black] 
    coordinates {
    (coremark,1.50)   [1.50]
    (coremark.ke,0.01) []
    (dhrystone,0.28)  [0.28]
    (enclave.ke,0.01) []
    (kyber,1.56)      [1.56]
    (kyber.ke,0.00)   []
    };

    % H (C) - crosshatch
    \addplot[pattern=crosshatch,pattern color=black] 
    coordinates {
    (coremark,4.60)   [4.60]
    (coremark.ke,0.02) []
    (dhrystone,0.71)  [0.71]
    (enclave.ke,0.00) []
    (kyber,4.71)      [4.71]
    (kyber.ke,0.00)   []
    };

    % H (M) - north east lines
    \addplot[pattern=north east lines,pattern color=black] 
    coordinates {
    (coremark,4.74)   [4.74]
    (coremark.ke,0.00) []
    (dhrystone,0.65)  [0.65]
    (enclave.ke,0.00) []
    (kyber,4.52)      [4.52]
    (kyber.ke,0.00)   []
    };

    % R² - crosshatch dots (dense)
    \addplot[pattern=crosshatch dots,pattern color=black] 
    coordinates {
    (coremark,0.030)   [0.030]
    (coremark.ke,0.000) []
    (dhrystone,0.005)  [0.005]
    (enclave.ke,0.000) []
    (kyber,0.036)      [0.036]
    (kyber.ke,0.000)   []
    };

    \legend{F (C),F (M),H (C),H (M),$R^2$}
    \end{axis}

    % Right axis for R² scaling
    \begin{axis}[
        hide axis,
        ymin=0, ymax=0.05,
        axis y line*=right,
        axis x line=none
    ]
    \end{axis}

    \end{tikzpicture}
    %\caption{ANOVA and Kruskal-Wallis results for hardware scaling effects. No statistically significant relationships found ($\alpha=0.05$)}
    %\label{fig:anova-kruskal-5pattern-wide-labels}
    \begin{tablenotes}
    \footnotesize
    \item F: ANOVA F-statistic; H: Kruskal–Wallis H-statistic; $R^2$: coefficient of determination.
    \item (C): CPU core scaling; (M): memory scaling.
    \item Kruskal–Wallis used due to potential non-normality in small-sample performance data\tnote{a}.
    \item[a] Kruskal–Wallis is robust to non-normal distributions and unequal variances.
    \end{tablenotes}
\end{threeparttable}
\caption{ANOVA and Kruskal–Wallis results for hardware scaling effects}
\label{fig:anova-kruskal-5pattern-wide-labels}
\end{figure}

Linear regression modeling (Table~\ref{tab:regression}) supports these observations, showing minimal slopes and $R^2$ values near zero for most workloads, indicating that neither core count nor memory size reliably predict performance changes. \texttt{coremark} exhibits modestly positive coefficients (+408 and +25 for cores and memory respectively), explaining some improvements seen in the scaling steps. In contrast, \texttt{dhrystone} and \texttt{kyber} show negative regression slopes, consistent with their observed performance degradation under scaling. Enclave variants uniformly display coefficients near zero, suggesting that enclave overheads may overshadow hardware scaling effects.

@@

% \begin{figure}[htbp]
% \centering
% \begin{tikzpicture}
% \begin{axis}[
%     ybar,
%     bar width=8pt,
%     width=14cm,
%     height=8cm,
%     ymin=-100, ymax=5500,
%     ylabel={Coefficient value},
%     symbolic x coords={coremark,coremark.ke,dhrystone,enclave.ke,kyber,kyber.ke},
%     xtick=data,
%     enlarge x limits=0.15,
%     ymajorgrids,
%     grid style=dashed,
%     legend style={at={(0.5,-0.20)},anchor=north,legend columns=3},
%     axis y line*=left,
%     axis x line*=bottom,
%     x tick label style={rotate=45,anchor=east},
%     nodes near coords,
%     every node near coord/.append style={font=\footnotesize, rotate=90, anchor=west}
% ]

% % Intercept – dots
% \addplot[pattern=dots,pattern color=black] 
% coordinates {
% (coremark,2852.203)
% (coremark.ke,5006.875)
% (dhrystone,1327.375)
% (enclave.ke,5046.875)
% (kyber,2847.875)
% (kyber.ke,5352.250)
% };

% % Cores Coef. – horizontal lines
% \addplot[pattern=horizontal lines,pattern color=black] 
% coordinates {
% (coremark,408.000)
% (coremark.ke,0.000)
% (dhrystone,-53.750)
% (enclave.ke,0.000)
% (kyber,-45.625)
% (kyber.ke,0.000)
% };

% % Memory Coef. – north east lines
% \addplot[pattern=north east lines,pattern color=black] 
% coordinates {
% (coremark,25.000)
% (coremark.ke,0.000)
% (dhrystone,-37.500)
% (enclave.ke,0.000)
% (kyber,-16.250)
% (kyber.ke,0.000)
% };

% \legend{Intercept,Cores Coef.,Memory Coef.}
% \end{axis}
% \end{tikzpicture}
% \caption{Linear regression coefficients of performance on hardware parameters. Positive values indicate performance gains; negative values indicate degradation.}
% \label{fig:regression-coefs-greyscale}
% \end{figure}

In summary, these findings highlight that hardware scaling alone is insufficient to guarantee improved performance in constrained embedded environments with enclave execution. Architectural and workload-specific characteristics critically modulate the impact of additional cores or memory. Without targeted architectural optimization or workload tuning, scaling resources may yield diminishing returns or even harm performance due to factors such as synchronization overheads, enclave context switches, or memory bandwidth limitations.

>>>>

Among the evaluated workloads, \texttt{coremark} stood out as the only one to exhibit consistent, positive scaling behavior. Performance improved by approximately 40.8\% with each additional core, and a similarly proportional uplift was observed with increased memory. This result is aligned with earlier observations: CoreMark’s mixed-instruction profile and parallelism-friendly design allow it to exploit both increased computational throughput and expanded memory capacity. As a compute-bound benchmark with moderate data locality, CoreMark scales predictably and efficiently with additional hardware provisioning.

@@

\section{Performance Sensitivity to Core and Memory Scaling}
\label{sec:hardware-impact}

Building on the preceding workload-specific performance analysis, this section investigates whether increasing CPU cores or available memory yields measurable performance improvements on our constrained RISC-V platform. Unlike traditional systems where scaling typically leads to linear or super-linear speedups, embedded systems with TEEs face distinct bottlenecks—such as synchronization overheads, memory bandwidth limits, and enclave management costs—that may inhibit or even reverse expected performance trends.

We systematically evaluated six representative workloads across multiple hardware configurations, incrementally increasing CPU core count and physical memory. The workloads include both native and enclave variants: \texttt{coremark}, \texttt{coremark.ke}, \texttt{dhrystone}, \texttt{enclave.ke}, \texttt{kyber}, and \texttt{kyber.ke}. Each was benchmarked across scaling steps, with performance tracked as relative throughput (operations per second).

Table~\ref{tab:hardware-impact} summarizes directional performance trends—whether performance improved, degraded, or remained constant—as well as the average relative performance change per scaling step.

@@

These results demonstrate a heterogeneous response to scaling: only one-third of workloads showed any improvement, while the majority exhibited no statistically meaningful gains. Notably, \texttt{coremark} and \texttt{coremark.ke} benefited moderately from increased resources, while other workloads—particularly those involving enclave transitions or cryptographic operations—remained largely invariant or slightly degraded under scaling.

To evaluate the robustness of these trends, we applied both parametric and non-parametric statistical tests. Specifically, one-way ANOVA tested for mean differences across scaling levels, while Kruskal–Wallis and Mann–Whitney U tests evaluated distributional differences without assuming normality. Additionally, Pearson correlation and linear regression were used to quantify trends and explained variance. Results are summarized in Table~\ref{tab:statistical-tests}.

\begin{table}[htbp]
\centering
\caption{Statistical test results for performance scaling across workloads. Parametric and non-parametric tests assess significance of changes with respect to CPU cores and memory.}
\label{tab:statistical-tests}
\begin{subtable}[t]{\textwidth}
\centering
\caption{Parametric Tests: ANOVA, Pearson correlation, and $R^2$.}
\vspace{0.5em}
\begin{tabular}{@{}l
                S[table-format=1.2] S[table-format=1.3]
                S[table-format=1.2] S[table-format=1.3]
                S[table-format=1.3]@{}}
\toprule
\textbf{Workload} 
& \textbf{F (Cores)} & \textbf{p (ANOVA)} 
& \textbf{r (Cores)} & \textbf{p (r)} 
& \textbf{$R^2$} \\
\midrule
\texttt{coremark}      & 1.50 & 0.218 & -0.16 & 0.052 & 0.030 \\
\texttt{coremark.ke}   & 0.01 & 0.975 &  0.00 & 0.999 & 0.000 \\
\texttt{dhrystone}     & 0.28 & 0.837 &  0.07 & 0.374 & 0.005 \\
\texttt{enclave.ke}    & 0.01 & 0.998 &  0.00 & 0.961 & 0.000 \\
\texttt{kyber}         & 1.56 & 0.202 & -0.17 & 0.053 & 0.036 \\
\texttt{kyber.ke}      & 0.00 & 1.000 &  0.00 & 1.000 & 0.000 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\footnotesize
\item Parametric tests assume normal distributions.
\item r: Pearson correlation; $R^2$: variance explained by linear regression.
\end{tablenotes}
\end{subtable}

\vspace{1.5em}

\begin{subtable}[t]{\textwidth}
\centering
\caption{Non-Parametric Tests: Kruskal–Wallis and Mann–Whitney U.}
\vspace{0.5em}
\begin{tabular}{@{}l
                S[table-format=2.2] S[table-format=1.3]
                S[table-format=2.2] S[table-format=1.3]@{}}
\toprule
\textbf{Workload} 
& \textbf{KW (H)} & \textbf{p (KW)} 
& \textbf{MWU (U)} & \textbf{p (MWU)} \\
\midrule
\texttt{coremark}      & 3.60 & 0.057 & 10.00 & 0.052 \\
\texttt{coremark.ke}   & 0.00 & 1.000 & 15.00 & 0.999 \\
\texttt{dhrystone}     & 0.20 & 0.654 & 14.00 & 0.374 \\
\texttt{enclave.ke}    & 0.01 & 0.961 & 15.00 & 0.961 \\
\texttt{kyber}         & 3.75 & 0.053 & 10.00 & 0.053 \\
\texttt{kyber.ke}      & 0.00 & 1.000 & 15.00 & 1.000 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\footnotesize
\item KW: Kruskal–Wallis H-statistic; MWU: Mann–Whitney U-statistic.
\item Non-parametric tests used due to small sample sizes and potential non-normality.
\end{tablenotes}
\end{subtable}
\end{table}

None of the statistical tests yielded significant $p$-values at the $\alpha = 0.05$ level, though \texttt{coremark} and \texttt{kyber} approached marginal significance in the Kruskal–Wallis and Mann–Whitney U tests. However, even in these cases, Pearson correlations and $R^2$ values indicate weak or negligible effect sizes. These results strongly suggest that, for most workloads, hardware scaling does not yield statistically meaningful performance improvements.

This interpretation is further supported by linear regression analysis in Table~\ref{tab:regression}, which reports slope coefficients and intercepts for each workload's performance as a function of core count and memory.

@@

Here again, \texttt{coremark} is the only workload to exhibit a positive response to hardware scaling, albeit with limited effect size. All enclave workloads display near-zero or flat regression slopes, underscoring the notion that scaling is generally ineffective when trusted execution overheads dominate the execution path.

In summary, while modest improvements are observed for select CPU-bound native workloads, scaling core or memory resources yields no consistent benefit—and occasionally even degrades performance—under enclave execution. These findings highlight the need for deeper architectural or runtime co-optimization if scaling is to deliver tangible performance benefits in secure embedded platforms.
